---
title: "InstacartEDA"
author: "Shivangi Vashi"
date: "5/26/2020"
output: html_document
---

Problem Statement
The Instacart data set is anonymized and contains samples of over 3 million grocery orders from 200,000+ Instacart users from Kaggle. 
The objective is to predict which previously purchased products (prior orders) will be in a user's next order (train and test orders) in order to explore the kinds of food Americans eat.

```{r}
library(dplyr)
library(tidyverse)
library(funModeling)
library(psych)
library(ggplot2)
library(plyr)

aisles<-read_csv(file.choose())
departments<-read_csv(file.choose())
order_products_prior<-read_csv(file.choose())
order_products_train<-read_csv(file.choose())
orders<-read_csv(file.choose())
products<-read_csv(file.choose())

aisles<-read_csv("/Users/shivangi/Documents/Analytics/Quarter 3/Data Mining/instacart-market-basket-analysis/aisles.csv")
departments<-read_csv("/Users/shivangi/Documents/Analytics/Quarter 3/Data Mining/instacart-market-basket-analysis/departments.csv")
order_products_prior<-read_csv("/Users/shivangi/Documents/Analytics/Quarter 3/Data Mining/instacart-market-basket-analysis/order_products__prior.csv")
order_products_train<-read_csv("/Users/shivangi/Documents/Analytics/Quarter 3/Data Mining/instacart-market-basket-analysis/order_products__train.csv")
orders<-read_csv("/Users/shivangi/Documents/Analytics/Quarter 3/Data Mining/instacart-market-basket-analysis/orders.csv")
products<-read_csv("/Users/shivangi/Documents/Analytics/Quarter 3/Data Mining/instacart-market-basket-analysis/products.csv")

```


### Create a new datafram named instacart by merging all datasets through inner join function, which only retains rows in all datasets.Here, we use order_products_prior dataframe includes only prior orders, so the new dataframe instacart will contain only these observations as well. 
```{r}
df<-list(aisles,departments,order_products_train,order_products_prior,orders,products)

instacart <- order_products_prior %>%
  inner_join(orders, by = c("order_id")) %>%
  inner_join(products, by = c("product_id")) %>%
  inner_join(aisles, by = c("aisle_id")) %>%
  inner_join(departments, by = c("department_id"))

summary(instacart)
```

### Pre-processing
```{r}

# Taking a quick look at the data
head(aisles)
head(departments)
head(order_products_prior)
head(order_products_train)
head(orders)
head(products)

#Remove missing values
aisles<-drop_na(aisles)
departments<-drop_na(departments)
order_products_prior<-drop_na(order_products_prior)
order_products_train<-drop_na(order_products_train)
orders<-drop_na(orders)
products<-drop_na(products)

dim(aisles)
dim(departments)
dim(order_products_prior)
dim(order_products_train)
dim(orders)
dim(products)

# Check if all classes ie data types is correct
sapply(aisles,class)
sapply(departments,class)
sapply(order_products_prior,class)
sapply(order_products_train,class)
sapply(orders,class)
orders<-orders %>%mutate(order_hour_of_day=as.numeric(order_hour_of_day))
orders
sapply(products,class)



```

### What did you do with the data in the context of exploration?
1. When do people order the most during the day? - shivangi
2. Frequency of reordering? - venus
3. Most popular products? - venus
4. Most popular department and most popular aisle? -rian

### What did you find? Why does that matter? 

### What would your proposed next steps be? 
- Apriori association rule algorithm
- classification- what are products are most likely to be reordered for each user?


### 1. When do people order the most during the day?

```{r}
pop_products<-inner_join(order_products_prior, products, by =c("product_id"))
pop_products<-as.data.frame(pop_products)
pop_products<-pop_products %>% select(5, 3)
top_products<-head(arrange(pop_products,desc(pop_products$add_to_cart_order)), n = 10)
top_products

ggplot(top_products)+
  geom_bar(aes(reorder(product_name,-add_to_cart_order),add_to_cart_order, fill= add_to_cart_order),
           stat = "identity", color = "grey", fill = "slateblue")+
  labs(title = "Most Popular Products Added to Cart Order", x = "product names", y = "add to cart order") +
  theme(axis.text.x = element_text(angle = 65, hjust = 1))

orders%>%
  group_by(order_hour_of_day)%>%
  summarise(Number_of_Orders=n())%>%
  ggplot(aes(y=Number_of_Orders, x=order_hour_of_day, fill=Number_of_Orders)) + geom_col()+ coord_cartesian(xlim = c(0, 24)) 

```
Clearly, most orders are placed during 9am to 5 pm.


### 2. What is the frequency of reordering? 
When exploring the frequency of reordering, first we filter prior eval_set and select data of days_since_prior_order, then we also clease all the na value from the dataset. 
We find out in this 3,000,000+ dataset, the average days of users who reorder their groceries is from 7 to 11 days. Many customers reorder after 7 days or 30 days. 
According to each customer's reordering behavior and how many days they would reorder through instacart,we can predict their next ordering day. Therefore, based on these analysis, it can provide suggestions to retails and suppliers which would be helpful for their purchase sales inventory stategies.
In the future, we still need to find out the reason why a lot customers reorder after 30 days and what products do they get then we can draw the conculsion. Also, we will find out the correlation among the frequency of reordering, order hour of day, and ordering products and tracking customer consuming behaviors.

```{r}
orders<-filter(orders,orders$eval_set=="prior")
days<-na.omit(orders$days_since_prior_order)
days<-as.data.frame(days)
summary(days)
describe(days)

ggplot(data=days, aes(x=days))+ 
  geom_boxplot(mapping=aes("var",days),colour="black",fill="lightblue")+
  xlab("")+
  ylab("days_since_prior_order")+
  ggtitle("Days since Prior Order Boxplot")

ggplot(data=days, aes(x=days))+ 
  geom_histogram(aes(y =..density..),stat = "bin", bins= 30, colour="black", fill="lavenderblush")+
  geom_density(alpha=.2,fill="plum")+
  ggtitle("Days since Prior Order Density Plot")

```


### 3. What are the most popular products?
By descending ordering products that are added into cart order, we find out the top 10 popular products out of 49,688 products as shown below. Surprisely, in our analysis, the top No.1 popular product is Organic 100% Grapefruit Juice.
It is important to find out what products people like to order and order the most, by doing so, we can have better prediction and preparation on the stock of these products. 
In our next steps, we will find out which departmants and aisles these products belong to, and how often do these top products reorder?

```{r}
pop_products<-inner_join(order_products_prior, products, by =c("product_id"))
pop_products<-as.data.frame(pop_products)
pop_products<-pop_products %>% select(5, 3)
top_products<-head(arrange(pop_products,desc(pop_products$add_to_cart_order)), n = 10)
top_products

ggplot(top_products)+
  geom_bar(aes(reorder(product_name,-add_to_cart_order),add_to_cart_order, fill= add_to_cart_order),
           stat = "identity", color = "grey", fill = "slateblue")+
  labs(title = "Most Popular Products Added to Cart Order", x = "product names", y = "add to cart order") +
  theme(axis.text.x = element_text(angle = 65, hjust = 1))

```
```{r}
install.packages("pacman")
pacman::p_load(dplyr, data.table, ggplot2, knitr, stringr, DT, treemap) 
library(data.table)
library(dplyr)
library(ggplot2)
library(knitr)
library(stringr)
library(DT)
library(treemap)

aisles <- fread('/Users/rian/Downloads/Instacart-Analysis-master/instacart-market-basket-analysis/aisles.csv')
departments <- fread('/Users/rian/Downloads/Instacart-Analysis-master/instacart-market-basket-analysis/departments.csv')
orders <- fread('/Users/rian/Downloads/Instacart-Analysis-master/instacart-market-basket-analysis/orders.csv')
order_products_train <- fread('/Users/rian/Downloads/Instacart-Analysis-master/instacart-market-basket-analysis/order_products__prior.csv')
order_products_prior <- fread('/Users/rian/Downloads/Instacart-Analysis-master/instacart-market-basket-analysis/order_products__prior.csv')
products <- fread('/Users/rian/Downloads/Instacart-Analysis-master/instacart-market-basket-analysis/products.csv')
```

```{r}
glimpse(aisles)
glimpse(departments)
glimpse(order_products_train)
glimpse(order_products_prior)
glimpse(orders)
glimpse(products)
```        

```{r}
#Data Preparation
#We recode and convert character variables to factors.
orders <- orders %>% mutate(order_hour_of_day = as.numeric(order_hour_of_day), eval_set = as.factor(eval_set))
products <- products %>% mutate(product_name = as.factor(product_name))
aisles <- aisles %>% mutate(aisle = as.factor(aisle))
departments <- departments %>% mutate(department = as.factor(department))
```

```{r}
#Data Preparation
tmp <- merge(orders, order_products_prior)
tmp2 <-  merge(tmp, products, by="product_id")
head(tmp2)
```

### Which Users Reorder?

To recommend the next item an user is most likely to add to his/ her cart, we have to understand how each group of users use the service. It becomes a classification problem that we can explore later, but for now we wanted to see how different users purchase products, how there can be correlation between their total number of orders, days since prior order, no. of items they add to their cart and their probability of reordering.   

```{r}
#user clusters
n_items_per_order <- tmp2 %>% group_by(order_id) %>% mutate(n_items=max(add_to_cart_order))
user <- n_items_per_order %>% 
  group_by(user_id) %>% 
  mutate(product_name = as.numeric(days_since_prior_order)) %>%
  summarize(total_orders = n_distinct(order_id), avg_days_since_prior_order= mean(days_since_prior_order, na.rm = TRUE), avg_no_items = mean(n_items), avg_reorder=mean(reordered))
head(user)
```

So what we can see from the table the people who have ordered more number of times, they have higher possibility of reordering the same products and they usually order after every three to four weeks. Which means that the products that gets ordered more will also have the higher probablity of being reordered. 
```{r}
order_products_train %>% 
  group_by(product_id) %>% 
  summarize(proportion_reordered = mean(reordered), n=n()) %>%
  ggplot(aes(x=n,y=proportion_reordered))+
  geom_point()+
  geom_smooth(color="red")+
  coord_cartesian(xlim=c(0,2000))

```

### Most popular department and most popular aisle?

So we wanted to see which are the most popular departments and aisles, as aparently these are the same department and aisles where will most reorder must occur. Later we can also find if which other product has the more probablity of being purchased with other. We can find that using association rule mining. To see the most popular departments and aisles we create a treemap where the size of the boxes shows the number of sales.

```{r}

tmp3 <- products %>% group_by(department_id, aisle_id) %>% summarize(n=n())
tmp3 <- tmp3 %>% left_join(departments,by="department_id")
tmp3 <- tmp3 %>% left_join(aisles,by="aisle_id")

product_portfolio <-order_products_train %>% 
  group_by(product_id) %>% 
  summarize(count=n()) %>% 
  left_join(products,by="product_id") %>% 
  ungroup() %>% 
  group_by(department_id,aisle_id) %>% 
  summarize(sumcount = sum(count)) %>% 
  left_join(tmp3, by = c("department_id", "aisle_id")) %>% 
  mutate(onesize = 1)

treemap(product_portfolio,index=c("department","aisle"),vSize="sumcount",title="",palette="Set3",border.col="#FFFFFF")
```

What we can see is produce, diary eggs are the most popular departments followed by snacks, pantry and others. But talking about departments, organic and non organic sector divides US consumers in a large way these days. So it will be interesting to see which products get most reordered. But at first we take a look at ration between the number of organic and non-organic foods.
```{r}
products <- products %>% 
    mutate(organic=ifelse(str_detect(str_to_lower(products$product_name),'organic'),"organic","not organic"), organic= as.factor(organic))
    
Organic_or_not <- order_products_train %>% 
  left_join(products, by="product_id") %>% 
  group_by(organic) %>% 
  summarize(count = n()) %>% 
  mutate(proportion = count/sum(count))
head(Organic_or_not)
Organic_or_not %>% 
  ggplot(aes(x=organic,y=count, fill=organic))+
  geom_bar(stat="identity")
```

So what we notice is the amount og non organic brands is way higher, as it is cheaper to produce and demand must be high as well. Then we try to find out which kind gets most reordered.

```{r}
reorder_organic_ornot <- order_products_train %>% left_join(products,by="product_id") %>% group_by(organic) %>% summarize(mean_reordered = mean(reordered))
head(reorder_organic_ornot)
reorder_organic_ornot %>% 
  ggplot(aes(x=organic,fill=organic,y=mean_reordered))+geom_bar(stat="identity")
```

What we get to see is organic foods are reordered in a much higher quantity. It could be because the people who active look for organic brands, they must have a much higher brand loyalty which makes it more prbable to be reordered.





