---
title: "InstacartEDA"
author: "Shivangi Vashi"
date: "5/26/2020"
output: html_document
---

Problem Statement
The Instacart data set is anonymized and contains samples of over 3 million grocery orders from 200,000+ Instacart users. The objective is to predict which previously purchased products (prior orders) will be in a user's next order (train and test orders) in order to explore the kinds of food Americans eat.

```{r}
library(dplyr)
library(tidyverse)

aisles<-read_csv(file.choose()) 
departments<-
order_products_prior



aisles<-read_csv("/Users/shivangi/Documents/Analytics/Quarter 3/Data Mining/instacart-market-basket-analysis/aisles.csv")
departments<-read_csv("/Users/shivangi/Documents/Analytics/Quarter 3/Data Mining/instacart-market-basket-analysis/departments.csv")
order_products_prior<-read_csv("/Users/shivangi/Documents/Analytics/Quarter 3/Data Mining/instacart-market-basket-analysis/order_products__prior.csv")
order_products_train<-read_csv("/Users/shivangi/Documents/Analytics/Quarter 3/Data Mining/instacart-market-basket-analysis/order_products__train.csv")
orders<-read_csv("/Users/shivangi/Documents/Analytics/Quarter 3/Data Mining/instacart-market-basket-analysis/orders.csv")
products<-read_csv("/Users/shivangi/Documents/Analytics/Quarter 3/Data Mining/instacart-market-basket-analysis/products.csv")

head(order_products_prior)
```

### What did you do with the data in the context of exploration?

1. When do people order the most during the day? - shivangi
2. Frequency of reordering? - venus
3. Most popular products? - venus
4. Most popular department and most popular aisle? -rian

### What did you find? Why does that matter? 

### What would your proposed next steps be? 
- Apriori association rule algorithm
- classification- what are products are most likely to be reordered for each user?
- 

```{r}

```

## 

```{r}

```

