---
title: "FinalProject"
author: "Shivangi Vashi"
date: "6/22/2020"
output: html_document
---

<center>
$\large Association~Rules~Mining$ </center>

<br><br>

For the Market Basket Analysis, only the order_products__prior and product are utilized and the same are joined on basis of product id. 
<br>


```{r}
library(plyr)
library(tidyverse)
library(data.table)
library(dplyr)
library(arules)
library(arulesViz)

order_products_prior<-fread("instacart-market-basket-analysis/order_products__prior.csv")
order_products_train<-fread("instacart-market-basket-analysis/order_products__train.csv")
orders<-fread("instacart-market-basket-analysis/orders.csv")
products<-fread("instacart-market-basket-analysis/products.csv")

set.seed(123)
user_fraction <- 0.1
users <- unique(orders$user_id)
sample_users <- sample(users, round(user_fraction * length(users)))

cat('Number of orders (before): ', nrow(orders))
orders <- orders[user_id %in% sample_users]
cat('Number of orders (after): ', nrow(orders))

# Training dataset
OrderProductPrior<-orders%>%inner_join(order_products_prior)
OrderProductPrior<-drop_na(OrderProductPrior)


#Testing dataset
OrderProductTrain<-orders%>%inner_join(order_products_train)
OrderProductTrain<-drop_na(OrderProductTrain)


dim(OrderProductPrior)
dim(OrderProductTrain)
head(OrderProductPrior)
head(OrderProductTrain)

```
<br>

Preparing data for apriori<br>
```{r}
basket_data <- left_join(OrderProductPrior, products, by='product_id')
basket_data <-  group_by(basket_data, order_id)
basket_data <- summarise(basket_data,items=as.vector(list(product_name)))
head(basket_data)


```

<br>
Since the eclat and apriori functions are only valid on transaction format, henceforth the format is converted to the transaction type with the help of below code snippet. 

```{r}
transactions=as(basket_data$items, 'transactions')
head(transactions)
length(transactions)
```

#### Setting the Support and Confidence Intervals<br>

Understanding support<br>

A value of Support =0.02 means that an item will be considered as frequent if at least 2 percent of all the baskets contain it.

Understanding Confidence<br>

Confidence is a measure of the strength of an association rule. It is the frequency of occurrence of the right-hand items in the rule from among those baskets that contain the items on the left-hand side of the rule.

So now we have created a function that will plot the number of rules we can generate depending on different support levels and varied upon different confidence levels. 
<br>
```{r}
# Support and confidence values
supportLevels <- c(0.1, 0.05, 0.01, 0.005)
confidenceLevels <- c(0.9, 0.8, 0.7, 0.6, 0.5, 0.4, 0.3, 0.2, 0.1)

# Empty integers 
rules_sup10 <- integer(length=9)
rules_sup5 <- integer(length=9)
rules_sup1 <- integer(length=9)
rules_sup0.5 <- integer(length=9)

# Apriori algorithm with a support level of 10%
for (i in 1:length(confidenceLevels)) {
  
  rules_sup10[i] <- length(apriori(transactions, parameter=list(sup=supportLevels[1], 
                                   conf=confidenceLevels[i], target="rules")))
  
}

# Apriori algorithm with a support level of 5%
for (i in 1:length(confidenceLevels)){
  
  rules_sup5[i] <- length(apriori(transactions, parameter=list(sup=supportLevels[2], 
                                  conf=confidenceLevels[i], target="rules")))
  
}

# Apriori algorithm with a support level of 1%
for (i in 1:length(confidenceLevels)){
  
  rules_sup1[i] <- length(apriori(transactions, parameter=list(sup=supportLevels[3], 
                                  conf=confidenceLevels[i], target="rules")))
  
}

# Apriori algorithm with a support level of 0.5%
for (i in 1:length(confidenceLevels)){
  
  rules_sup0.5[i] <- length(apriori(transactions, parameter=list(sup=supportLevels[4], 
                                    conf=confidenceLevels[i], target="rules")))
  
}
```




```{r}
# Number of rules found with a support level of 10%
plot1 <- qplot(confidenceLevels, rules_sup10, geom=c("point", "line"), 
               xlab="Confidence level", ylab="Number of rules found", 
               main="Apriori with a support level of 10%") +
  theme_bw()

# Number of rules found with a support level of 5%
plot2 <- qplot(confidenceLevels, rules_sup5, geom=c("point", "line"), 
               xlab="Confidence level", ylab="Number of rules found", 
               main="Apriori with a support level of 5%") + 
  scale_y_continuous(breaks=seq(0, 10, 2)) +
  theme_bw()

# Number of rules found with a support level of 1%
plot3 <- qplot(confidenceLevels, rules_sup1, geom=c("point", "line"), 
               xlab="Confidence level", ylab="Number of rules found", 
               main="Apriori with a support level of 1%") + 
  scale_y_continuous(breaks=seq(0, 50, 10)) +
  theme_bw()

# Number of rules found with a support level of 0.5%
plot4 <- qplot(confidenceLevels, rules_sup0.5, geom=c("point", "line"), 
               xlab="Confidence level", ylab="Number of rules found", 
               main="Apriori with a support level of 0.5%") + 
  scale_y_continuous(breaks=seq(0, 130, 20)) +
  theme_bw()

# Subplot
plot1 
plot2 
plot3
plot4
```


```{r}
rules1 <- apriori(transactions, parameter = list(supp = 0.1, conf = 0.5, maxlen=3), control = list(verbose = FALSE))
as(rules1,"data.frame")
rules2 <- apriori(transactions, parameter = list(supp = 0.001, conf = 0.4, maxlen=3), control = list(verbose = FALSE))
as(rules2,"data.frame")
rules3 <- apriori(transactions, parameter = list(supp = 0.005, conf = 0.1, maxlen=3), control = list(verbose = FALSE))
as(rules3,"data.frame")

plot(rules2, method="paracoord",  control=list(alpha=.5, reorder=TRUE))


```
<br><br>

